---
title: "ilex-physcraper-res"
output: rmarkdown::html_document
---

### The Physcraper tag

T1281-M2478

### Input files

Physcraper rewrites input files for a couple reasons: reproducibility, taxon name matching, and taxon reconciliation.
It writes the config file down if none was provided

### Run files

Files in here are also automatically written down by physcraper.

blast runs, alignments, raxml trees, bootstrap

The trees are reconstructed using RAxML, with tip labels corresponding to local ids (e.g., otu42009, otuPS1) and not taxon names (e.g., *Helwingia japonica*), nor taxonomic ids (e.g., ott or ncbi). Branch lengths are proportional to relative substitution rates.
The RAxML tree with taxon names as tip labels is saved on the `outputs_tag` folder.

```{r pars2, include=FALSE}
par(mar=c(0.5,1,1,0.5), xpd=NA)
```
```{r updated-tree-otus, echo=TRUE, out.width='100%', fig.height= 5, fig.retina=3, dpi=300, out.extra='style="background-color: #6B8E23; padding:10px; display: inline-block;"'}
updated_tree_otus <- ape::read.tree(file = "../data/ilex-local/run_T1281-M2478/RAxML_bestTree.2020-06-29")
ape::plot.phylo(ape::ladderize(updated_tree_otus), type = "fan", cex = 0.25, label.offset = 0.01, edge.width = 0.5)
ape::add.scale.bar(cex = 0.3, font = 1, col = "black")
mtext("Updated tree - otu tags as labels", side = 3, line = -1)
```

### Output files

A variety of files are automatically written down by physcraper.

A nexson tree with all types of tip labels is saved in here.
From this tree, a tree with any kind of label can be produced.
By default, the updtaed tree with taxon names as tip labels is saved
in the `output_tag` folder as `updated_taxonname.tre`.


```{r read-updated-tree-taxonname, echo=TRUE}
updated_tree_taxonname <- ape::read.tree(
  file = "../data/ilex-local/outputs_T1281-M2478/updated_taxonname.tre")
# ape::plot.phylo(ape::ladderize(updated_tree_taxonname), cex = 0.35)
updated_tree_taxonname$tip.label %in% c("Helwingia_japonica_otu420083", "Helwingia_chinensis_otu420099")

updated_tree <- ape::read.tree(file = "../data/pg_2827_tree6577/run_pg_2827tree6577_run4/RAxML_bestTree.2020-07-31")
updated_drop <- ape::drop.tip(updated_tree, c("otu420083", "otu420099"))

```
```{r pars3, include=FALSE}
par(mar=c(0,1,2,0), xpd=NA)  # adjust margins for all plots after this?
```
```{r updated-tree-taxonname, echo=TRUE, out.width='100%', fig.height= 5, fig.retina=3, dpi=300,  out.extra='style="background-color: #9ecff7; padding:10px; display: inline-block;"'}
ape::plot.phylo(ape::ladderize(updated_tree_taxonname), type = "fan", cex = 0.25, label.offset = 0.01, edge.width = 0.5)
ape::add.scale.bar(cex = 0.3, font = 1, col = "black")
mtext("Updated tree - Taxon names as labels  ", side = 3, line = -1)
```
