---
title: "ilex-physcraper-test"
output: rmarkdown::html_document
---

To perform a test of Physcraper functionality, we pruned the Gottlieb et al. tree down ~20%, excluding the outgroups, corresponding to approximately 9 tips. We then performed a Physcraper run to test if we would recover the dropped tips.
The input alignment will be automatically trimmed y Physcraper

We performed the following steps:

1. [Pruning the tree](#pruning)
1. [Matching the names](#matching)
1. [Running Physcraper](#running)
1. [Checking the results](#checking)
1. Analyzing Physcraper results

# Pruning the Gottlieb et al. tree

### Getting the tree from the Open Tree of Life database

```{r, echo = TRUE}
tree6577_pg_2827 <- rotl::get_study_tree(study_id = "pg_2827", tree_id = "tree6577", tip_label = "ott_taxon_name")

```

### Identifying the outgroups
```{r, echo=TRUE}
outgroup_index <- match(c("Helwingia_chinensis", "Helwingia_japonica"), tree6577_pg_2827$tip.label)
```


### Getting the ingroup tip labels
```{r, echo=TRUE}
ingroup_tip.label <- tree6577_pg_2827$tip.label[-outgroup_index]
```


### Calculating the number of tips to keep
```{r, echo=TRUE}
ingroup_ntip <- length(ingroup_tip.label)
ingroup_ntip* 0.8
```

### Choosing at random the ingroup tips to drop
```{r, echo = TRUE}

set.seed(19022021)
todrop <- sample(ingroup_tip.label, ingroup_ntip-37)
todrop
```

### Dropping the tips from the tree

```{r, echo=TRUE}
pruned_tree <- ape::drop.tip(phy = tree6577_pg_2827, tip = todrop)
pruned_tree
```

### Saving the tree as newick and nexus

```{r, echo = TRUE}
ape::write.tree(phy = pruned_tree, file = "../data-raw/ilex-test/tree6577_pg_2827_pruned.tre")
ape::write.nexus(pruned_tree, file = "../data-raw/ilex-test/tree6577_pg_2827_pruned.nex")
```

# Matching the tree tip labels

### First generate a simple text file containing one tip label per line

```{r, echo = TRUE}

write(paste(pruned_tree$tip.label, collapse = "\n"), file = "../data-raw/ilex-test/pruned_tree_tip_labels.txt")
```
# Running the test

### Locally

```
physcraper_run.py -tf physcraperex/data-raw/ilex-test/tree6577_pg_2827_pruned.tre -tfs newick -a physcraperex/data-raw/alignments/T1281-M2478.nex -as nexus -ti physcraperex/data-raw/ilex-test/pruned_tree_tip_labels_mapped/main.json -db local_blast_db -nt 8 -o physcraperex/data/ilex-test-local
```

### Remotely
```
physcraper_run.py -tf physcraperex/data-raw/ilex-test/tree6577_pg_2827_pruned.tre -tfs newick -a physcraperex/data-raw/alignments/T1281-M2478.nex -as nexus -ti physcraperex/data-raw/ilex-test/pruned_tree_tip_labels_mapped/main.json -o physcraperex/data/ilex-test-remote
```


# Checking the results of the test

### Getting some information from the otu_info table

```{r, echo=TRUE, warning= FALSE}
otu_info_T1281_M2478 <- readr::read_delim("../data/ilex-test-local/inputs_T1281-M2478/otu_info_T1281-M2478.csv", "\t", escape_double = FALSE, trim_ws = TRUE)
DT::datatable(otu_info_T1281_M2478)
```

```{r}
# Get the number of tips in input tree that are not in alignment
sum(!otu_info_T1281_M2478$`^physcraper:in_current_aln`)

# Get the taxon names of those tips
not_in_ali <- otu_info_T1281_M2478$`^ot:ottTaxonName`[!otu_info_T1281_M2478$`^physcraper:in_current_aln`]
```

There are 6 tips on tree that are not in the alignment, corresponding to the following taxa: 

`r paste0("\n- _", not_in_ali, "_", collapse = "\n")`

Physcraper prunes these tips form the input tree, leaving it with 39 - 6 = 33 tips:

```{r, echo = TRUE}
taxonname_tre <- ape::read.tree(file = "../data/ilex-test-local/inputs_T1281-M2478/taxonname.tre")

taxonname_tre

physcraper_T1281_M2478_tre <- ape::read.tree(file = "../data/ilex-test-local/inputs_T1281-M2478/physcraper_T1281-M2478.tre")

physcraper_T1281_M2478_tre
```


### Checking that the original alignment was appropriately trimmed

```{r, echo=TRUE}
taxonname_fas <- ape::read.dna(file = "../data/ilex-test-local/inputs_T1281-M2478/taxonname.fas", format = "fasta")
taxonname_fas

physcraper_T1281_M2478_fas <- ape::read.dna(file = "../data/ilex-test-local/inputs_T1281-M2478/physcraper_T1281-M2478.fas", format = "fasta")

physcraper_T1281_M2478_fas
```

```{r, echo = TRUE}
geiger::name.check(phy = taxonname_tre, data = taxonname_fas)
geiger::name.check(phy = physcraper_T1281_M2478_tre, data = physcraper_T1281_M2478_fas)
```
