---
title: "ilex-physcraper-test"
output: rmarkdown::html_document
---

To perform a test of Physcraper functionality, we pruned the Gottlieb et al. tree down ~20%, excluding the outgroups, corresponding to 9 tips. We then performed a Physcraper run to test if we would recover the dropped tips.
The input alignment will NOT be automatically trimmed by Physcraper

We performed the following steps:

1. [Pruning the tree](#pruning)
1. [Matching the names](#matching)
1. [Running Physcraper](#running)
1. [Checking the results](#checking)
1. Analyzing the results

# Pruning the Gottlieb et al. tree

### Getting the tree from the Open Tree of Life database

```{r, echo = TRUE}
tree6577_pg_2827 <- rotl::get_study_tree(study_id = "pg_2827", tree_id = "tree6577", tip_label = "original_label")
tree6577_pg_2827$tip.label
```

### Identifying the outgroups
```{r, echo=TRUE}
outgroup_index <- match(c("Helwingia_chinensis", "Helwingia_japonica"), tree6577_pg_2827$tip.label)
```


### Getting the ingroup tip labels
```{r, echo=TRUE}
ingroup_tip.label <- tree6577_pg_2827$tip.label[-outgroup_index]
```


### Calculating the number of tips to keep
```{r, echo=TRUE}
ingroup_ntip <- length(ingroup_tip.label)
ingroup_ntip* 0.8
```

### Choosing at random the ingroup tips to drop
```{r, echo = TRUE}

set.seed(19022021)
todrop <- sample(ingroup_tip.label, ingroup_ntip-37)
todrop
```

### Dropping the tips from the tree

```{r, echo=TRUE}
pruned_tree <- ape::drop.tip(phy = tree6577_pg_2827, tip = todrop)
pruned_tree
```

### Saving the tree as newick and nexus

```{r, echo = TRUE, eval = FALSE}
ape::write.tree(phy = pruned_tree, file = "../data-raw/ilex-test/tree6577_pg_2827_pruned.tre")
ape::write.nexus(pruned_tree, file = "../data-raw/ilex-test/tree6577_pg_2827_pruned.nex")
```

# Matching the tree tip labels

### First generate a simple text file containing one tip label per line

```{r, echo = TRUE}

write(paste(pruned_tree$tip.label, collapse = "\n"), file = "../data-raw/ilex-test/pruned_tree_tip_labels.txt")
```

### Go to OpenTree's TNRS matching bulk tool

Available at this link https://tree.opentreeoflife.org/curator/tnrs/



# Pruning the alignment

### Read the alignment

```{r, echo=TRUE}
ali_T1281_M2478 <- ape::read.nexus.data(file = "../data-raw/alignments/T1281-M2478.nex")
names(ali_T1281_M2478)
```

### Prune it

```{r, echo=TRUE}
in_ali <- !names(ali_T1281_M2478) %in% todrop
ali_T1281_M2478_pruned <- ali_T1281_M2478[in_ali]
```

### Save the pruned alignment

```{r, echo = TRUE, eval = FALSE}
ape::write.nexus.data(x = ali_T1281_M2478_pruned, file = "../data-raw/ilex-test/alignment_T1281_M2478_pruned.nex", interleaved = FALSE)

# dendropy.dataio.nexusreader.InvalidCharacterStateSymbolError: Error parsing data source 'physcraperex/data-raw/ilex-test/alignment_T1281_M2478_pruned.nex' on line 7 at column 33: Unrecognized character state symbol for state alphabet 'DNA' (DnaStateAlphabet) : '/'
```

### Modify the slash added when multistate characters are present in the matrix

I did this by hand, by looking for "/" symbol for example in "atcg/t-tct" and removing it and enclosing the adjacent bases within curly braces to indicate multistate character, "atc{gt}-tct"
I found some cases with three different states, for example "atc/g/t-tct", would be "at{cgt}-tct"

# Running the test

### Locally

```
physcraper_run.py -tf physcraperex/data-raw/ilex-test/tree6577_pg_2827_pruned.tre -tfs newick -a physcraperex/data-raw/ilex-test/alignment_T1281_M2478_pruned.nex -as nexus -ti physcraperex/data-raw/ilex-test/pruned_tree_tip_labels_mapped/main.json -db /absolute/path/local_blast_db -nt 8 -o physcraperex/data/ilex-test-local-1
```
Trying it with the otu_info.csv file from a previous run:

```
physcraper_run.py -tf physcraperex/data-raw/ilex-test/tree6577_pg_2827_pruned.tre -tfs newick -a physcraperex/data-raw/alignments/T1281-M2478.nex -as nexus -ti physcraperex/data/ilex-local/run_T1281-M2478/otu_info_T1281-M2478.json -db /absolute/path/local_blast_db -nt 8 -o physcraperex/data/ilex-test-local
```

### Remotely
```
physcraper_run.py -tf physcraperex/data-raw/ilex-test/tree6577_pg_2827_pruned.tre -tfs newick -a physcraperex/data-raw/alignments/T1281-M2478.nex -as nexus -ti physcraperex/data-raw/ilex-test/pruned_tree_tip_labels_mapped/main.json -o physcraperex/data/ilex-test-remote
```


# Checking the results of the test

### Getting information from the OTU info table

```{r, echo=TRUE, warning= FALSE}
otu_info_T1281_M2478 <- readr::read_delim("../data/ilex-test-local-1/inputs_alignment_T1281_M2478_pruned/otu_info_alignment_T1281_M2478_pruned.csv", "\t", escape_double = FALSE, trim_ws = TRUE)
DT::datatable(otu_info_T1281_M2478)
```

```{r}
# Verify that all tips in input tree are in alignment
sum(!otu_info_T1281_M2478$`^physcraper:in_current_aln`)
```


### Looking at the output OTU info file

```{r, echo=TRUE, warning= FALSE}

otu_info_alignment_T1281_M2478_pruned <- readr::read_delim("../data/ilex-test-local-1/outputs_alignment_T1281_M2478_pruned/otu_info_alignment_T1281_M2478_pruned.csv", "\t", escape_double = FALSE, trim_ws = TRUE)
DT::datatable(otu_info_alignment_T1281_M2478_pruned)
```

### Getting the OTT ids of the dropped taxa

```{r, echo = TRUE}

todrop_tnrs <- rotl::tnrs_match_names(names = todrop)
```

### Looking for the dropped taxa in the OTU info table

```{r, echo = TRUE}

# Using the OTT id:
todrop_tnrs$ott_id %in% otu_info_alignment_T1281_M2478_pruned$`^ot:ottId`

# Using the unique taxon name:
todrop_tnrs$unique_name %in% otu_info_alignment_T1281_M2478_pruned$`^ot:ottTaxonName`

# Getting the taxa in original tree that are not in the updated tree
notinupdated <- todrop_tnrs$unique_name[!todrop_tnrs$ott_id %in% otu_info_alignment_T1281_M2478_pruned$`^ot:ottId`]

```
The following taxa were in original tree but were not recovered with this Physcraper run:

`r paste0("- _", notinupdated, "_", collapse = "\n ")`


### Why does this happen?

We looked for the NCBI GenBank accession numbers reported in the [supplementary material of the original publication](https://bsapubs.onlinelibrary.wiley.com/action/downloadSupplement?doi=10.3732%2Fajb.92.2.352&file=ajb20352-sup-0002.pdf) for
the ITS sequences of each of the missed taxa,
_Ilex warburgii_, _I. dimorphophylla_ and _I. percoriacea_. The first too have updated the originally reported accession numbers from U92600/U92601 and U92592/U92593 to a single accession number [AH007153](https://www.ncbi.nlm.nih.gov/nuccore/AH007153.2?expand-gaps=on) and
[AH007149](https://www.ncbi.nlm.nih.gov/nuccore/AH007149), respectively.
For _I. percoriacea_, the NCBI accession number reported in the priginal alignment is [AH007156](https://www.ncbi.nlm.nih.gov/nuccore/AH007156), and it remains the same (as of `r format(Sys.time(), "%a %b %d, %Y")`).
Inspecting the sequences, all three contain a middle gap of 100 undetermined nucleotide bases (Ns). These Ns were not present in these sequences in the original alignment. Physcraper drops these sequence because they are unalignable without any manual curation.









Note: updated_taxonname.tree and labelled_tag.tre are exactly the same tree
